\documentclass{subfiles}
\begin{document}\label{sec:RB-delete}
    Similarly for insertion, deleting a node from a red-black tree works as for the BSTs,
        with a few differences depending on the original position of the node to remove.
        As before, we make use of an auxiliary procedure to fix any any violations.
        In \autoref{proc:RB-delete} the pseudo code to implement deletion is shown,
        while in \autoref{proc:RB-delete-fixup} we show the pseudo code for the auxiliary procedure.

    Let us consider which properties may be violated: first,
        property 2 fails to hold when the node to remove was the root and the new root is black,
        Property 4 fails when both the replacing node and it's right child are red;
        lastly, properties 1 and 5 fail when nodes are moved around.

    \subfile{../TikZ/Procedure 4 - RB delete}

    We can analyze \autoref{proc:RB-delete-fixup} as eight cases, pairwise symmetric.
        We consider for this reason only case 1-4.
        To ease our explanation, let \(z\) be the node to remove,\(z.p\) its parent, 
        \(y\) theminimum in the \(z\)'s right sub-tree and let \(x\) be \(y\)'s right child.
        \begin{description}
            \item [Case 1: \(\mathbf{x}\)'s sibling \(\mathbf{w}\) is ``Red''.]
                We know that \(w\) must have a black children,
                thus a rotation on \(x.p\) will solve the violation.

            \item [Case 2: \(\mathbf{w}\) is ``Black'' and so are its children.]
                Since both \(x\) and \(w\) are black, we simply ``push'' the blackness
                upward, and repeat the procedure on the parent.

            \item [Case 3: \(\mathbf{w}\) is ``Black'', its left child is ``Red'' and the left child is ``Red''.]
                A simple color exchange between \(w\) and \(w.left\) followed by a right rotation,
                fixes the issue.

            \item [Case 4: \(\mathbf{w}\) is ``Black'' and its right child is ``Red''.]
                Some color exchanges and a left rotation on \(x.p\) will solve the violations.
        \end{description}
        As for insertion, once the loop has finished the only property that may still fail is property
            2, which is fixed by recolorig the root.

        \subfile{../TikZ/Procedure 5 - RB delete fixup}
        
\end{document}
